---
title: "Explore Obesity Data"
subtitle: "DA5020 / Assignment A3"
author: "Chen, Angeline & Hobbs, Tyrian"
date: "`r Sys.Date()`"
output: pdf_document
---

## Setup & Import 

```{r Q1 setup & import}
# Load libraries & tibble
library(dplyr)
library(tidyr) #for the replace.na function
library(knitr)

df.full <- read.csv("http://artificium.us/datasets/ObesityPredictionData-Modified.csv") |>
  tibble::as_tibble()
glimpse(df.full)
```
## Cleaning & Missing Values
```{r Q2 cleaning}
# Removing unnecessary columns and filling empty values
df.clean <- df.full %>% select(-CH2O, -SCC, -FAF, -TUE, -CALC)

df.clean <- df.clean |>
  group_by(Gender) |>
  mutate(
    Height = replace_na(Height, mean(Height, trim = 0.05, na.rm = TRUE)),
    Age = replace_na(Age, median(Age, na.rm = TRUE))
  ) |>
  ungroup()

df.clean |> summarise(across(everything(), ~sum(is.na(.))))
```
After removing unnecessary columns and replacing the missing columns in Height and Age, there are no more missing values in our dataset.

## Outliers & BMI Feature
``` {r Q3 outliers}
#Removing outliers, calculating BMI
df.clean <- df.clean |> 
  #using abs() to remove both tails at once
  filter(abs(Weight - mean(Weight, na.rm = TRUE)) <= 2.5 * sd(Weight, na.rm = TRUE)) |> 
  mutate(BMI = Weight/(Height**2))
```

We filtered `r nrow(df.full)-nrow(df.clean)` rows with weights outside 2.5x the standard deviation. Afterwards, the mean BMI is `r round(mean(df.clean$BMI), digits=3)`.

## Sampling & Height Comparison
```{r Q4 sampling}
# Getting the partial dataset
set.seed(44376)
df.partial <- df.clean |> sample_frac(0.4)

# Calculating heights
mean_height_full <- mean(df.clean$Height, na.rm = TRUE)
CI_height_full <- t.test(df.clean$Height, conf.level = 0.95)$conf.int

mean_height_part <- mean(df.partial$Height, na.rm = TRUE)
CI_height_part <- t.test(df.partial$Height, conf.level = 0.95)$conf.int
```

The full dataset has a mean height of `r round(mean_height_full, digits=3)` and 95% confidence interval [`r round(CI_height_full, digits=3)`]. The partial dataset has mean height `r round(mean_height_part, digits=3)` and 95% confidence interval [`r round(CI_height_part, digits=3)`]. Both the means and confidence intervals are incredibly similar, which is encouraging; in theory a random sample should have the same mean as the population it draws from. 

## Summaries & Correlation
``` {r Q5 summaries}
# Getting patient numbers and gender percentages
overview <- df.clean |>
  summarise(
    Total = n(),
    PercentFemale = sum(Gender == "Female", na.rm = TRUE) / n() * 100,
    PercentMale = sum(Gender == "Male", na.rm = TRUE) / n() * 100
  )
```

There are `r overview$Total` total patients, of which `r round(overview$PercentFemale, digits=3)`% are female and `r round(overview$PercentMale, digits=3)`% are male.

``` {r Q5 table}
#Make & Display the counts table
obesity_table <- df.clean |>
  filter(Obesity %in% c("Obesity_Type_I", "Obesity_Type_II")) |>
  group_by(Gender, Obesity) |>
  summarise(n = n(), .groups = "drop_last")

obesity_table |> kable(
  col.names = c("Gender", "Obesity Type", "Count"),
  booktabs = TRUE,
  caption = "Type I and Type II obesity counts by gender")
```

``` {r Q5 pearson}
# Finding correlation
cor <- cor.test(df.clean$Age, df.clean$BMI, method = "pearson")
```
Age and BMI have a correlation coefficient of `r round(cor$estimate, digits=3)`. This is positive, meaning older people tend to have higher BMI, but close to zero, meaning the correlation is fairly weak. The p-value is `r cor$p.value`, well under 0.05, meaning the correlation is statistically significant.